---
title: "Lizard Lengths in the Jornada Basin"
author: "Jacob Weverka"
date: "12/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)


```

In this analysis I will compare body lengths of females from two lizard species sampled from the Jornada Basin: the Checkered Whiptail (*Cnemidophorus tesselatus*) and the Desert Grassland Whiptail (*C. uniparens*). I will examine the mean and standard deviation of the body length for each species, then use a Welch Two Sample t-test and a Cohen's d effect size test to understand whether females of these species have a different mean size.

```{r}
# Read in data


lizard_data = read_csv(here::here("data/lizards.csv"))

lizards = lizard_data %>%
  select(spp, sex, total_length)


```


```{r}

# Select only females 

females = lizards %>%
  filter(spp == "CNUN" | spp == "CNTE",
         sex == "F")
```


```{r include=FALSE}

# Exploratory vizualizations of data looking for normality - don't include in final report!

ggplot(females, aes(x = total_length)) +
  geom_histogram() +
  facet_wrap(. ~ spp)

ggplot(females, aes(sample = total_length)) +
  geom_qq() +
  facet_wrap(. ~ spp)

# Data look fairly normal (mayble slightly deviated from normal)


```

```{r}
length_sum = females %>%
  group_by(spp) %>%
  summarise(mean_length = mean(total_length, na.rm = TRUE), sd_length = sd(total_length, na.rm = TRUE), n = n())


length_sum_round = length_sum %>%
  mutate(mean_length = round(mean_length, 2),
         sd_length = round(sd_length, 2))


knitr::kable(length_sum_round,
             col.names = c("Species", "Mean Length", "SD Length", "Sample Size"),
             caption = "Means and standard deviations of body lengths for female Checkered Whiptails and Desert Grassland Whiptails")
```




```{r fig.cap = "Body lengths of  female Checkered Whiptails and Desert Grassland Whiptails. Mean and standard deviation of body lengths for each species are shown in blue, while individual lizard body lengths are in grey."}
# Vizualize means, SEs and actual data

ggplot() +
  ggbeeswarm::geom_beeswarm(data = females, aes(x = spp, y = total_length),
                            size = 3,
                            alpha = 0.1) +
  geom_point(data = length_sum, aes(x = spp, y = mean_length),
             color = "slateblue4",
             size = 5) +
  geom_errorbar(data = length_sum,
                aes(x = spp,
                    ymin = mean_length - sd_length,
                    ymax = mean_length + sd_length),
                color = "slateblue4",
                width = 0.2) +
  xlab("Species") +
  ylab("Body Length (mm)") +
  scale_x_discrete(name = "Species",
                   labels = c("CNTE" = "Checkered Whiptail", "CNUN" = "Desert Grassland Whiptail")) +
  theme_bw()
```






```{r}
# Now a t-test

cnun = females %>%
  filter(spp == "CNUN") %>%
  pull(total_length)

cnte = females %>%
  filter(spp == "CNTE") %>%
  pull(total_length)

species.t = t.test(cnte, cnun)

species.tidy = broom::tidy(species.t)

```

```{r}
# And a cohen's d effect size

species.d = effsize::cohen.d(cnte, cnun, na.rm = TRUE)


```

Female Checkered Whiptails were one average `r round(species.tidy$estimate, 2)` mm longer than female Desert Grassland Whiptails (t(`r round(species.tidy$parameter, 2)`) = `r round(species.tidy$statistic, 2)`, p < `r format(signif(species.tidy$p.value, 3), scientific = TRUE)`). With a Cohen's d statistic of `r round(species.d$estimate, 2)` we can be confident that there is a fairly large difference in mean body length between the two species.



Data Citation:
[Lightfoot, D. 2017. Lizard pitfall trap data (LTER-II, LTER-III) ver 36. Environmental Data Initiative.](https://doi.org/10.6073/pasta/ff37b4cace16a9943575f3fd7067064e)








