---
title: "Political and Environmentalist Self-Identification"
author: "Jacob Weverka"
date: "12/6/2020"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(janitor)

```


```{r}

#Reading in the data and select the columns we need.

survey = read_csv(here::here("data/conservation_survey.csv"))

simple = survey %>%
  select(ENV_ID, POL_OR)

```

In this analysis, I will examine survey results to determine whether political ideology is related to the likelihood of individuals self-identifying as environmentalists. Here, I'm binning respondents into political categories based on the question:  "How do you self-identify politically?" Responses with values 1-3 are conservative, responses with values 5-7 are liberal. I will not be including other responses in this analysis.

```{r}
binned = simple %>%
  mutate(category = 
           case_when(
              POL_OR <= 3 ~ "Conservative",
              POL_OR >= 5 ~ "Liberal"
              ),
         response = 
           case_when(
             ENV_ID == 1 ~ "Yes",
             ENV_ID == 2 ~ "No",
             ENV_ID == 3 ~ "Unsure"
           )
  ) %>%
  drop_na()
```





```{r}
# Here is is a table showing the survey results for liberals and conservatives asked if they agree with the statement "I consider myself an environmentalist."


political_table = binned %>%
  tabyl(category, response)

political_p = political_table %>%
  adorn_percentages() %>%
  adorn_pct_formatting(digits = 2)

political_proportions = political_p %>%
  adorn_ns() %>%
  relocate(category, Yes, No, Unsure) %>%
  clean_names("upper_camel")

lib_yes = political_p %>%
  filter(category == "Liberal") %>%
  select(Yes)

con_yes = political_p %>%
  filter(category == "Conservative") %>%
  select(Yes)





knitr::kable(political_proportions, caption = "Proportions of Liberals and Conservatives that Identify as Environmentalists. Actual survey counts are in parentheses next to percentages.")

```



.
```{r}

# This code, now converts these same data into contingency table format so we can do a $\chi$^2^ test

env_ct = political_table %>%
  column_to_rownames(var = "category")


```




```{r}

# Now, performing a $\chi$^2^ test.

env_x2 = chisq.test(env_ct)


env_tidy = broom::tidy(env_x2)

```
There is a significant association between political preference and self-identification as an environmentalist ($\chi$^2^(`r env_tidy$parameter`). = `r round(env_tidy$statistic,2)`, *p* = `r round(env_tidy$p.value, 3)`). It appears that politically liberal people are more likely to identify as environmentalists, with `r lib_yes ` identifying as environmentalists, compared to `r con_yes` of conservatives self-identifying as environmentalists.        `



Data citation: [Andrews Forest LTER Site and C. Batavia. 2019](https://doi.org/10.6073/pasta/926e6270e324a1322a900da14d38b96c)




























